---
title: "Modeling directly from antibody levels"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Modeling directly from antibody levels}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, output=FALSE}
library(serosv)
```

## Mixture model

**Proposed model**

Two-component mixture model for test result $Z$ with $Z_j (j = \{I, S\})$ being the latent mixing component having density $f_j(z_j|\theta_j)$ and with $\pi_{\text{TRUE}}(a)$ being the age-dependent mixing probability can be represented as

$$
f(z|z_I, z_S,a) = (1-\pi_{\text{TRUE}}(a))f_S(z_S|\theta_S)+\pi_{\text{TRUE}}(a)f_I(z_I|\theta_I)
$$

The mean $E(Z|a)$ thus equals

$$
\mu(a) = (1-\pi_{\text{TRUE}}(a))\mu_S+\pi_{\text{TRUE}}(a)\mu_I$$

From which the true prevalence can be calculated by

$$
\pi_{\text{TRUE}}(a) = \frac{\mu(a) - \mu_S}{\mu_I - \mu_S}
$$

Force of infection can then be calculated by

$$
\lambda_{TRUE} = \frac{\mu'(a)}{\mu_I - \mu(a)}
$$

**Fitting data**

To fit the mixture data, use `mixture_model` function

```{r}
df <- vzv_be_2001_2003[vzv_be_2001_2003$age < 40.5,]
df <- df[order(df$age),]
data <- df$VZVmIUml
model <- mixture_model(antibody_level = data)
model$info
```

```{r}
plot(model)
```

sero-prevalence and FOI can then be esimated using function `estimate_from_mixture`

```{r}
est_mixture <- estimate_from_mixture(df$age, data, mixture_model = model, threshold_status = df$seropositive, sp=83, monotonize = FALSE)
plot(est_mixture)
```
